\textbf{Proof Reconstruction in Classical Propositional Logic}

### 2017-05-09 slides/slides.tex
[ slides ] added more footnotes.

\only<2->{\node[right = 0.8cm of problem, hyperlink node=tptp-syntax](tptp){
\includegraphics[scale=0.7]{figures/tptp}
};
}
\only<3->{
\node[right= 0.8cm of tptp, hyperlink node=metis] (metis)
  {\includegraphics[scale=0.7]{figures/metis}\,
  \footnote{\url{http://www.gilith.com/software/metis}}
  };
}
\only<4->{
\node[right= 0.8cm of metis, hyperlink node=tstp-syntax] (tstp)
  {\includegraphics[scale=0.7]{figures/tstp}};
}
\only<5->{
\node[below= 1.5cm of tstp, hyperlink node=athena] (athena)
  {\includegraphics[scale=0.7]{figures/athena}
  \footnote{\url{http://github.com/jonaprieto/athena}}
  };
}
\only<6->{
\node[left = 0.6cm of athena, hyperlink node=agda-file] (agdafile)
  {\includegraphics[scale=0.7]{figures/agdafile}
  };
}
\only<7->{
\node[left = 0.5cm of agdafile, hyperlink node=agda] (agda)
  {\includegraphics[scale=0.7]{figures/agda}
  \footnote{\url{http://github.com/agda/agda}}
  };
}
\only<8->{
\node[below = 0.18cm of problem, hyperlink node=verified-example] (verified)
  {\includegraphics[scale=0.7]{figures/verified}};
}
\only<9->{
\node[below = 0.18cm of verified, opacity=0.5, hyperlink node=failure-example] (failure)
  {\includegraphics[scale=0.7]{figures/failure}};
}

@@

\begin{frame}[fragile, label=reconstruction]{Proof Reconstruction: Overview}
\vfill
\begin{tikzpicture}
\node(problem) at (0,0)
   {\includegraphics[scale=0.7]{figures/problem}};
\node[right = 0.8cm of problem, hyperlink node=tptp-syntax](tptp){
\includegraphics[scale=0.7]{figures/tptp}
};
\node[right= 0.8cm of tptp, hyperlink node=metis] (metis)
  {\includegraphics[scale=0.7]{figures/metis}};
\node[right= 0.8cm of metis, hyperlink node=tstp-syntax] (tstp)
  {\includegraphics[scale=0.7]{figures/tstp}};
\node[below= 1.5cm of tstp, hyperlink node=athena] (athena)
  {\includegraphics[scale=0.7]{figures/athena}};
\node[left = 0.6cm of athena, hyperlink node=agda-file] (agdafile)
  {\includegraphics[scale=0.7]{figures/agdafile}};
\node[left = 0.5cm of agdafile, hyperlink node=agda] (agda)
  {\includegraphics[scale=0.7]{figures/agda}};
\node[below = 0.18cm of problem, hyperlink node=verified-example] (verified)
  {\includegraphics[scale=0.7]{figures/verified}};
\node[below = 0.18cm of verified, opacity=0.5, hyperlink node=failure-example] (failure)
  {\includegraphics[scale=0.7]{figures/failure}};

@@

\begin{itemize}
    \item Parsing of \texttt{TSTP} language
    \item Creation and analysis of \hyperlink{tstp-dag}{\textbf{DAG}} derivations
    \item Analysis of inference rules used in the \texttt{TSTP} derivation
    \item \texttt{Agda} code generation
\begin{table}[!ht]
\begin{center}
\scalebox{0.8}{
{\renewcommand{\arraystretch}{1.3}%
\begin{tabular}{|ll| }
\hline
\textbf{Library} & \textbf{Purpose} \\
\hline
\texttt{\hyperlink{agda-prop}{\texttt{Agda-Prop}}}& axioms and theorems of Classical Propositional Logic\\
\texttt{\hyperlink{agda-metis}{\texttt{Agda-Metis}}} & versions of the inference rules used by \texttt{Metis}\\
\hline
\end{tabular}}
}
\end{center}
\label{tab:library}
\end{table}

@@

@@

proof : Γ ⊢ goal
proof =
  ⇒-elim
    atp-splitGoal
    (∧-intro proof₀ proof₁)
\end{minted}
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=failure-example]{Bug in the Printing of the Proof}
  {\texttt{Metis}' v2.3 (release 20161108) - Issue\footnote{\url{https://github.com/gilith/metis/issues/2}}}
% {}
\begin{tptp}
$ cat problem.tptp
fof(goal, conjecture,
  ((p <=> q) <=> r) <=> (p <=> (q <=> r))).
\end{tptp}
\begin{tptp}
$ metis --show proof problem.tptp
...
fof(normalize_2_0, plain,
  (~ p & (~ q <=> ~ r) & (~ p <=> (~ q <=> ~ r))),
  inference(canonicalize, [], [negate_2_0])).
fof(normalize_2_1, plain, ~ p <=> (~ q <=> ~ r),
    inference(conjunct, [], [normalize_2_0])).
fof(normalize_2_2, plain, ~ q <=> ~ r,
    inference(conjunct, [], [normalize_2_0])).
fof(normalize_2_3, plain, ~ p,
  inference(conjunct, [], [normalize_2_0])).
fof(normalize_2_4, plain, $false,
    inference(simplify, [],
      [normalize_2_1, normalize_2_2, normalize_2_3])).
...
\end{tptp}
\end{frame}

@@

\RightLabel{\scriptsize{simplify}}
\TrinaryInfC{$\bot$}
\end{prooftree}
\pause
The bug was caused by the conversion of \texttt{Xor} sets to \texttt{Iff} lists.
After reporting this, Joe fixed the printing of \texttt{canonicalize} inference rule
\[ φ := ¬ p ∧ (¬ q ⇔ ¬ r) ∧ (¬ p ⇔ (¬ q ⇔ {\color{red} r}))\]

\end{frame}

@@

\only<1>{At the moment, the communication between \texttt{Agda} and
the ATPs is unidirectional because the ATPs are being used as oracles \citep{Sicard2015}
\vfill
}
\only<1>{
\inputminted{cagda}{Or.agda}
}

@@

\only<7->{ \node[left= 2.75cm of atp, rectangle, text=white,fill=black, align=left, font=\tiny] (proved)
{
\texttt{\$ agda Or.agda}\\
\texttt{\$ apia Or.agda --atp=online-metis}\\
\texttt{Metis---2.3 proved the conjecture}
  }
};

@@

\begin{frame}[label=pending-work]{Pending Work}
\begin{itemize}
\item There are missing cases with the \texttt{simplify} inference
\item Is not clear, how  \texttt{canonicalize} inference choose what
normal form use to transform the formulas
\item Splitting a goal in a list of subgoals is not verified yet
\end{itemize}
\end{frame}

\begin{frame}[label=future-work]{Contributions and Future Work}
\vfill
\begin{table}[!ht]
\begin{center}
\scalebox{0.8}{
{\renewcommand{\arraystretch}{1.3}%
\begin{tabular}{|ll|}
\hline
\textbf{Name} & \textbf{Purpose} \\
\hline
\href{http://github.com/jonaprieto/agda-prop}{\texttt{Agda-Prop}}     & lib. for syntax and theorems of Clasical Propositional Logic\\
\href{http://github.com/jonaprieto/agda-metis}{\texttt{Agda-Metis}}   & lib. for versions of the inference rules used by \texttt{Metis}\\
\href{http://github.com/jonaprieto/athena}{\texttt{Athena}}           & soft. Translator in Haskell for \texttt{Metis'} \texttt{TSTP} files to \texttt{Agda}\\
\href{http://github.com/jonaprieto/online-atps}{\texttt{OnlineATPs}}  & soft. A client in Haskell for \texttt{SystemOnTPTP} of \texttt{TPTP} World\\[1ex]
\hline
\end{tabular}}
}
\end{center}
\label{tab:library}
\end{table}
\vfill

@@

\begin{frame}[fragile,label=metis-inference-rules]{\texttt{Metis}' Inference Rules}
\vfill
\begin{prooftree}
\AxiomC{}
\RightLabel{axiom}
\UnaryInfC{$C$}
\end{prooftree}
\vfill
\begin{prooftree}
\AxiomC{}
\RightLabel{assume $L$}
\UnaryInfC{$L ∨ ¬ L$}
\end{prooftree}
\vfill
\begin{prooftree}
\AxiomC{$C$}
\RightLabel{subst $σ$}
\UnaryInfC{$σC$}
\end{prooftree}
\vfill
\begin{prooftree}
\AxiomC{$L ∨ C$}
\AxiomC{$¬ L ∨ D$}
\RightLabel{resolve $L$}
\BinaryInfC{$C ∨ D$}
\end{prooftree}
\vfill
\begin{prooftree}
\AxiomC{}
\RightLabel{refl $t$}
\UnaryInfC{$t = t$}
\end{prooftree}
\vfill
\begin{prooftree}
\AxiomC{}
\RightLabel{eq $L$ $p$ $t$}
\UnaryInfC{$¬ (L[p] = t) ∨ ¬ L ∨ L[ p ↦ t]$}
\end{prooftree}
\vfill
\end{frame}
 \begin{tikzpicture}[overlay, remember picture, scale=0.5]
    \tikzset{shift={(current page.center)}, xshift=11.8cm,yshift=-8.4cm}
    \node[fill=plum,text=white, hyperlink node = metis]
      (go) at (0,0) {\tiny Go Back};
  \end{tikzpicture}
\end{document}

>>>>

\only<2->{\node[right = 0.8cm of problem, hyperlink node=tptp-syntax](tptp){
\includegraphics[scale=0.7]{figures/tptp}
};
}
\only<3->{
\node[right= 0.8cm of tptp, hyperlink node=metis] (metis)
  {\includegraphics[scale=0.7]{figures/metis}\,
  \footnote{\url{http://www.gilith.com/software/metis}}
  };
}
\only<4->{
\node[right= 0.8cm of metis, hyperlink node=tstp-syntax] (tstp)
  {\includegraphics[scale=0.7]{figures/tstp}};
}
\only<5->{
\node[below= 1.5cm of tstp, hyperlink node=athena] (athena)
  {\includegraphics[scale=0.7]{figures/athena}
  \footnote{\url{http://github.com/jonaprieto/athena}}
  };
}
\only<6->{
\node[left = 0.6cm of athena, hyperlink node=agda-file] (agdafile)
  {\includegraphics[scale=0.7]{figures/agdafile}
  };
}
\only<7->{
\node[left = 0.5cm of agdafile] (agda)
  {\includegraphics[scale=0.7]{figures/agda}
  \footnote{\url{http://github.com/agda/agda}}
  };
}
\only<8->{
\node[below = 0.18cm of problem, hyperlink node=verified-example] (verified)
  {\includegraphics[scale=0.7]{figures/verified}};
}
\only<9->{
\node[below = 0.18cm of verified, opacity=0.5, hyperlink node=failure-example] (failure)
  {\includegraphics[scale=0.7]{figures/failure}};
}

@@

\begin{frame}[fragile, label=reconstruction]{Proof Reconstruction: Overview}
\vfill
\begin{tikzpicture}
\node(problem) at (0,0)
   {\includegraphics[scale=0.7]{figures/problem}};
\node[right = 0.8cm of problem, hyperlink node=tptp-syntax](tptp){
\includegraphics[scale=0.7]{figures/tptp}
};
\node[right= 0.8cm of tptp, hyperlink node=metis] (metis)
  {\includegraphics[scale=0.7]{figures/metis}};
\node[right= 0.8cm of metis, hyperlink node=tstp-syntax] (tstp)
  {\includegraphics[scale=0.7]{figures/tstp}};
\node[below= 1.5cm of tstp, hyperlink node=athena] (athena)
  {\includegraphics[scale=0.7]{figures/athena}};
\node[left = 0.6cm of athena, hyperlink node=agda-file] (agdafile)
  {\includegraphics[scale=0.7]{figures/agdafile}};
\node[left = 0.5cm of agdafile] (agda)
  {\includegraphics[scale=0.7]{figures/agda}};
\node[below = 0.18cm of problem, hyperlink node=verified-example] (verified)
  {\includegraphics[scale=0.7]{figures/verified}};
\node[below = 0.18cm of verified, opacity=0.5, hyperlink node=failure-example] (failure)
  {\includegraphics[scale=0.7]{figures/failure}};

@@

\begin{itemize}
    \item Parsing of \texttt{TSTP} language\footnote{\label{parsing}\url{https://github.com/agomezl/tstp2agda}}\textsuperscript{,}\footnote{\url{https://github.com/jonaprieto/tstp2agda}}
    \item Creation\textsuperscript{\ref{parsing}} and analysis of \hyperlink{tstp-dag}{\textbf{DAG}} derivations
    \item Analysis of inference rules used in the \texttt{TSTP} derivation
    \item \texttt{Agda} code generation
\begin{table}[!ht]
\begin{center}
\scalebox{0.8}{
{\renewcommand{\arraystretch}{1.3}%
\begin{tabular}{|ll| }
\hline
\textbf{Library} & \textbf{Purpose} \\
\hline
\texttt{\hyperlink{agda-prop}{\texttt{Agda-Prop}}}& axioms and theorems of Classical Propositional Logic\\
\texttt{\hyperlink{agda-metis}{\texttt{Agda-Metis}}} & versions of the inference rules used by \texttt{Metis}\\
\hline
\end{tabular}}
}
\end{center}
\label{tab:library}
\end{table}

@@

Γ = [ a ]

@@

proof : Γ ⊢ goal
proof =
  ⇒-elim
    atp-splitGoal            -- q ∧ (q ⇒ p) ⇒ p
    (∧-intro proof₀ proof₁)
\end{minted}
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=failure-example]{Bug\footnote{\url{https://github.com/gilith/metis/issues/2}} in the Printing of the Proof}
  {\texttt{Metis}' v2.3 (release 20161108)}
% {}
\begin{tptp}
$ cat problem.tptp
fof(goal, conjecture,
  ((p <=> q) <=> r) <=> (p <=> (q <=> r))).
\end{tptp}
\begin{tptp}
$ metis --show proof problem.tptp
...
fof(normalize_2_0, plain,
  (~ p & (~ q <=> ~ r) & (~ p <=> (~ q <=> ~ r))),
  inference(canonicalize, [], [negate_2_0])).
fof(normalize_2_1, plain, ~ p <=> (~ q <=> ~ r),
    inference(conjunct, [], [normalize_2_0])).
fof(normalize_2_2, plain, ~ q <=> ~ r,
    inference(conjunct, [], [normalize_2_0])).
fof(normalize_2_3, plain, ~ p,
  inference(conjunct, [], [normalize_2_0])).
fof(normalize_2_4, plain, $false,
    inference(simplify, [],
      [normalize_2_1, normalize_2_2, normalize_2_3])).
...
\end{tptp}
\end{frame}

@@

\RightLabel{\scriptsize{simplify}}
\TrinaryInfC{$\bot$}
\end{prooftree}
\pause
The bug was caused by the conversion of \texttt{Xor} sets to \texttt{Iff} lists.
After reporting this, Joe fixed the printing of \texttt{canonicalize} inference rule
\[ φ := ¬ p ∧ (¬ q ⇔ ¬ r) ∧ (¬ p ⇔ (¬ q ⇔ {\color{red} r}))\]
\gobackreconstruction
\end{frame}

@@

\only<1>{At the moment, the communication between \texttt{Agda} and
the ATPs is unidirectional because the ATPs are being used as oracles \citep{Sicard2015}
\vfill
}
\only<1>{
\inputminted{cagda}{Or.agda}
\begin{tikzpicture}[overlay, remember picture, scale=0.5]
\tikzset{shift={(current page.center)}, xshift=6cm,yshift=-2cm}
\node(agdafile) at (0,0){
  \includegraphics[scale=0.8]{figures/agda-pragmas}};
\end{tikzpicture}
}

@@

\only<7->{ \node[left= 3cm of atp, rectangle, text=white,fill=black, align=left, font=\tiny] (proved)
{
\texttt{\$ agda Or.agda}\\
\texttt{\$ apia Or.agda --atp=online-metis}\\
\texttt{Metis---2.3 proved the conjecture}
  }
};

@@

\begin{frame}[label=pending-work]{Pending Work}
\begin{itemize}
\item Complete implementation for \texttt{simplify} inference\footnote{\url{https://github.com/gilith/metis/issues/3}}
\item Complete implementation for \texttt{canonicalize} inference: what
normal form use to transform the formulas
\item Complete implementation for Splitting a goal in a list of subgoals
\end{itemize}
\end{frame}

\begin{frame}[label=future-work]{Contributions and Future Work}
\vfill
\begin{table}[!ht]
\begin{center}
\scalebox{0.8}{
{\renewcommand{\arraystretch}{1.3}%
\begin{tabular}{|ll|}
\hline
\textbf{Name} & \textbf{Purpose} \\
\hline
\href{http://github.com/jonaprieto/agda-metis}{\texttt{Agda-Metis}}   & Implementation of \texttt{Metis}' inference rules\\
\href{http://github.com/jonaprieto/agda-prop}{\texttt{Agda-Prop}}     & Syntax and theorems of Classical Propositional Logic\\
\href{http://github.com/jonaprieto/athena}{\texttt{Athena}}           & Translator for \texttt{Metis'} \texttt{TSTP} files to \texttt{Agda}\\
\href{http://github.com/jonaprieto/online-atps}{\texttt{OnlineATPs}}  & Client to use ATPs from \texttt{SystemOnTPTP} of \texttt{TPTP} World\\
\href{http://github.com/jonaprieto/prop-pack}{\texttt{Prop-Pack}}     & Collection of \texttt{TPTP} problems to test \texttt{Athena}\\[1ex]
\hline
\end{tabular}}
}
\end{center}
\label{tab:library}
\end{table}
\vfill

@@

\begin{frame}[fragile,label=metis-inference-rules]{\texttt{Metis}' Inference Rules}
\vfill
\begin{prooftree}
\AxiomC{}
\RightLabel{axiom}
\UnaryInfC{$C$}
\end{prooftree}
\vfill
\begin{prooftree}
\AxiomC{}
\RightLabel{assume $L$}
\UnaryInfC{$L ∨ ¬ L$}
\end{prooftree}
\vfill
\begin{prooftree}
\AxiomC{$C$}
\RightLabel{subst $σ$}
\UnaryInfC{$σC$}
\end{prooftree}
\vfill
\begin{prooftree}
\AxiomC{$L ∨ C$}
\AxiomC{$¬ L ∨ D$}
\RightLabel{resolve $L$}
\BinaryInfC{$C ∨ D$}
\end{prooftree}
\vfill
\begin{prooftree}
\AxiomC{}
\RightLabel{refl $t$}
\UnaryInfC{$t = t$}
\end{prooftree}
\vfill
\begin{prooftree}
\AxiomC{}
\RightLabel{eq $L$ $p$ $t$}
\UnaryInfC{$¬ (L[p] = t) ∨ ¬ L ∨ L[ p ↦ t]$}
\end{prooftree}
\vfill
 \begin{tikzpicture}[overlay, remember picture, scale=0.5]
    \tikzset{shift={(current page.center)}, xshift=11.8cm,yshift=-8.4cm}
    \node[fill=plum,text=white, hyperlink node = metis]
      (go) at (0,0) {\tiny Go Back};
  \end{tikzpicture}
\end{frame}
\end{document}
